Enriching Object Detection with 2D-3D Registration and Continuous Viewpoint

### 2014-11-13 egpaper_final.tex
excised intro into separate file

%%%%%%%%% TITLE
\title{Enrich Object Detection : 2D-3D registration and continuous viewpoint estimation}

@@

%%%%%%%%% BODY TEXT
\section{Introduction}

\begin{figure}[t]
  \centering
  % \includegraphics[width=0.9\linewidth]{schematics} 
  \includegraphics[width=0.99\linewidth]{front} % \\[-5pt]
  % (a)\\[0pt]
  % \setlength\tabcolsep{0pt}
  % \begin{tabular}{cc}
  %     \includegraphics[width=0.45\linewidth]{car/orig2} & 
  %     \includegraphics[width=0.45\linewidth]{car/overlay2}
  %  %   \begin{tabular}{cc}
  %  %       \includegraphics[width=0.16\linewidth]{cad_car/2001-2004 Ford Ranger Edge} & 
  %  %       \includegraphics[width=0.16\linewidth]{cad_car/2002 Dodge Ram FedEx} \\ 
  %  %       \includegraphics[width=0.16\linewidth]{cad_car/2007-Nissan-Versa-_-Tiida-SL} & 
  %  %       \includegraphics[width=0.16\linewidth]{cad_car/2008-Jeep-Cherokee} \\ 
  %  %       \includegraphics[width=0.16\linewidth]{cad_car/2009 Toyota Cargo} & 
  %  %       \includegraphics[width=0.16\linewidth]{cad_car/2010-Chevrolet-Aveo5-LT} \\ 
  %  %       \includegraphics[width=0.16\linewidth]{cad_car/2012-Citroen-DS4}  
  %  %   \end{tabular}
  % \end{tabular}\\
  % (b)
  \caption{Using a bank of CAD models, we generate NZ-WHO templates which can be used to either detect objects directly or initialize pose of an object in the bounding box from other detectors. Once initialization is given, we use NZ-WHO to propose and validate plausible pose on-the-fly and further tune the translation, 3D viewpoint and focal length continuously.}
  % \vspace{-1em}
  \label{fig:front}
\end{figure}

% Object detection task has long been about predicting rectangular bounding boxes around the object. In this workthere has been significant many powerful object detectors such as Convolutional Neural Network or Part-Based Models \cite{Girshick14, Felzenszwalb10, Pepik12}

Matching 3D CAD model to an image has been a classical problem that dates back to eighties \cite{Lowe87}. Lowe \etal , the vision community instead focused on the object detection tasks that estimates a rectangular bounding box around an object\cite{Felzenszwalb10, Girshick14}. To infer what is more than a rectangle, several attempts have been made to estimate the pose of an object \cite{Xiang12, Pepik12, Fidler12, Hejrati14} or even finding corresponding CAD model as well \cite{Aubry14, Lim14}.

However, there has been increasing demand for accurate pose and 2D-3D matching. Notably, in graphics community, there has been growing demand for accurate 2D-3D matching \cite{Kholgade14, Chen13, Kostas14} for image manipulation, editing and data augmentation respectively.

Recently, with the powerful computing capabilities and a large collection of CAD models, the 2D-3D matching (registration) problem gets more and more approachable but due to inherent difficulty and limited performance of existing methods, many applications still require supervision to achieve 2D-3D matching.

% In medical community, registering 3D scan data to 2D image enabled various minimally invasive surgeries \cite{Markelj12}.

For category-level matching, \cite{Xiang12, Hejrati14, Pepik12} used 3D parts or aspect to represent underlying category-level 3D geometry of the object. With the advent of large public CAD datasets \cite{Trimble, Yobi}, the instance-level object detection and 2D-3D registration became possible. \cite{Aubry14, Lim14} used large collection of CAD models to faithfully align 3D CAD model to an image by rendering CAD models and creating 2D templates for various viewpoints.

Data-driven instance-level 2D-3D matching algorithms such as \cite{Aubry14, Lim14} require a large number of rendering and 2D templates for each of the CAD model with specific viewpoint. Since there are so many templates that it becomes almost impossible to get all the true positive image from training data. Instead \cite{Aubry14, Lim14} used whitening as a preprocessing stage \cite{Hariharan12} to make templates more discriminative. Then calibration is done on a small training dataset.

Although whitening to create template is faster than some of instance-level matching algorithms such as training Exemplar-SVM \cite{Malisiewicz11} or complicated models such as LEVAN \cite{Divvala14}, whitening and calibration large number of templates takes very long time. Also, once the fixed discrete viewpoint templates have been made and calibrated, it is difficult to fine-tune the prediction made using the large collection of templates.

%% Our contribution
With this growing usage of WHO template and Computer Graphics rendering, we decided to analyze WHO feature
[Find More Usage of WHO and CG rendering \cite{Pepik12}]

In our paper, we present Non-Zero Whiten Histograms of Orientations (NZ-WHO) to dramatically speed up the template generation and get rid of time consuming calibration and yet achieves performance on par with calibrated templates. First, we will present NZ-WHO is more discriminative than WHO and introduce Conjugate Gradient method to speed up the whitening and create a high resolution template in just 80 milliseconds. Also, using this fast NZ-WHO template generation, we present \textbf{on-the-fly} template generation to fine tune an initial pose estimation to continuously refine translation, viewpoints (yaw, roll, pitch) and focal length.

Our pipeline is simple, fast only requires CAD models and no validation images for calibration. This can be used for post processing stage of generic object detection. Our pipeline requires only several minutes to make templates for rough viewpoints and

We propose a simple method to register the most similar CAD model to an image and fine-tune viewpoint continuously without discriminative part search, training nor calibration. Our method is simple, fast and light-weight in terms of memory so that it can be easily applied on top of generic object detectors as a final stage of detection or can be used as a standalone detector.

% Although our method is geared toward instance-level 2D-3D matching, we evaluated the performance of the templates on PASCAL 2007 and PASCAL 2012 and PASCAL 3D dataset to measure the performance.

Our tuning stage not only refine position, scale and viewpoint, but also focal length as well to literally "fine-tune" the initial proposal. Though our 2D-3D matching method can be used as a rudimentary object detector, our simple tuning stage can be applied to any generic object detector to achieve state-of-the-art viewpoint estimation method.

Finally, we show several applications using our method. Since we have metadata such as CAD model and geometry, we reconstruct the object using 2D-3D matching. Also, since we do not need training image, we show that our 2D-3D matching performs better than standard detectors in categories where training data lacks. We then use our method for image query to pull out closest CAD model in a bank of CAD models.

Finally, unlike several CAD model alignment methods that uses  WHO \cite{Aubry13, Aubry14, Lim14}, we propose a novel pipeline that does not require calibration which is infeasible in continuous viewpoint estimation. We propose a simple add-on stage that can enrich object detections with high-quality 2D-3D alignment. The pipeline is simple and light weight so this can be easily adapted to be used for any generic object detectors. Also to facilitate the future work, all our alignment result on PASCAL12 and code are publicly available online.

% Recently, matching a 3D CAD model to an image, 2D-3D matching (registration), is gaining great momentum in various fields of study. 2D-3D matching using Structure from Motion \cite{Sattler11} uses existing 3D model to match the model to a query image. Also image manipulation using 3D CAD models has proven its potential. Kholgade \etal \cite{Kholgade14} and Chen \etal \cite{Chen13} require human input to match 3D models to an image to created realistic morphing or manipulation of the registered object. However, all these application requies human input. In this paper, we propose a novel on-the-fly template generation algorithm that can generate template quickly and validate 2D-3D pose and focal length continuously.

% Our code has been made publicly online \cite{Choy14} so that our algorithm can be used as a pipeline for 2D-3D registration after general purpose object detectors.

\subsection{Approach Overview}

Our method is in line with \cite{Aubry14, Lim14} in the sense that we use only renderings not real images. However, rather than using fixed sized mid-level patches, we use whole object and make root templates. By using root templates, there is no need to learn a discriminative patch dictionary or to calibrate the patches and an additional inference stage is not required. However, whitening the whole template is prohibitive for several reasons. All templates have different aspect ratios making it difficult to generate the covariance matrix $\Sigma$. Since the root templates are larger than mid-level patches, a very large auto-covariance $\Gamma$ is required. Lastly, time complexity of decorrelating a template with $n$ cells increases as $O(n^3)$, resulting in slow template generation and large residuals due to numerical errors accumulated while decomposing the covariance matrix $\Sigma$.

To overcome these difficulties, we propose the Non-Zero Whitened Histograms of Orientations with iterative Conjugate Gradient to generate templates in real time and enrich rectangular boxes with high quality 2D-3D matching. The 2D-3D matching provides a CAD model, sub-category, rendering, depth, viewpoint and focal length.

Our method can work as a standalone 2D-3D matching algorithm, or even as a standalone ensemble of exemplar-based object detectors (Figure \ref{fig:front}) but the real power comes from combining generic object detectors and complement rectangular bounding boxes. Our method is simple in design; is several orders of magnitude faster than traditional method; requires only CAD models; doesn't need images nor calibration, yet is powerful in giving high-quality metadata. These properties make our method ideal for complementing object detections.

To complement detection bounding boxes, we first search possible 2D-3D matching by running cached discrete viewpoint templates. Then, for the most probable match, we refine the initial matching using Metropolis Hastings. Since it requires iterations of proposal and validation, we present real time analysis of NZ-WHO template generation to show that this can be done on-the-fly. The fine tuning stage process jointly optimizes translation, yaw, roll, pitch, scale, and focal length continuously and also searches for different CAD models.

>>>>

%%%%%%%%% TITLE
\title{Enriching Object Detection with 2D-3D Registration and Continuous Viewpoint}
%\title{Enrich Object Detection : 2D-3D registration and continuous viewpoint estimation}

@@

%%%%%%%%% BODY TEXT
\section{Introduction}
\label{sec:intro}
\input{intro}
